### 2017-01-10 draft.tex
LaTeX files are added

% Suha's Proposal
\documentclass{article}

\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{arydshln}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{color}

\newtheorem{dfn}{Definition}
\newtheorem{lem}{Lemma}

\title{Existence of Forward Simulations for Particular Data Structures}
\begin{document}
\maketitle
\input{prelim}
\input{queue}
\section{Existence of Forward Simulations for Stack Implementations that have Fixed Pop Linearization Points}
\input{commit}
\input{set}
\end{document}