### 2017-01-12 draftCAV/refProofs.tex

%!TEX root = draft.tex

\section{Refinement Proofs}

Refinement between two libraries $L_1$ and $L_2$ (and thus, linearizability) can be proved using simulation relations which roughly, require that 
the library $L_2$ can mimic every step of $L_1$ using a (possibly empty) sequence of steps. Mainly, there are two kinds of simulation
relations, forward or backward depending on whether the preservation of steps is proved starting from a similar state
forward or backward. In general, forward simulations are easier to establish, but they are incomplete, i.e., refinement may hold while
there exists no forward simulation. On the other hand, proofs based on backward simulations are more difficult, but modulo the
addition of history variables (auxiliary variables recording events in the implementation) backward simulations are complete for 
proving refinement~\footnote{Adding history variables has no impact on the incompleteness of forward simulations.}.

This section introduces notions used to enlarge the scope of proofs based on forward simulations, which is the main goal of this paper.
We first introduce a stronger notion of refinement that compares the set of traces of two libraries projected
on a larger alphabet of actions than the set of call/return actions. These actions are called \emph{observable actions},
and they may represent for instance, linearization points which informally, are points in time where an operation 
is conceptually effectuated. We then define a notion of forward simulation which is complete for proving
this notion of refinement provided that the alphabet of the ``abstract'' library $L_2$ equals the set of 
observable actions and $L_2$ is deterministic. We discuss an application of this result, namely,
proving the correctness of libraries with ``fixed linearization points'', where the so-called ``linearization points''
are fixed to particular implementation actions.

\begin{dfn}\label{def:gref}
Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. 
The library $L_1$ \emph{$\Gamma$-refines} $L_2$ if{f} $Tr(L_1) | \Gamma \subseteq Tr(L_2) | \Gamma$.
%We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
%for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
\end{dfn}

Notice that $\Gamma$-refinement implies refinement for any $\Gamma$ as in Definition~\ref{def:gref}.

We define a notion of \emph{forward} simulation that can be used to prove $\Gamma$-refinement. 
A dual notion of \emph{backward} simulation is defined in Appendix~\ref{app:backSim}. 
%We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement. 
For a relation $R\subseteq A\times B$, $R[X]$ denotes the set of elements of $B$ related to elements of $X$ by $R$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.

%We can extend the relations between libraries by introducing simulation relations. We will later show that simulation relations imply refinement.

\begin{dfn}
Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
\begin{itemize}
\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$ 
\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0<i<|@s|$.
%$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0\leq i<|@s| -1$.
% where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $\delta_2(u,\gamma, u')$. 
\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma_1\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0\leq i<|@s|$.  
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}
%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
A $\Gamma$-forward simulation shows that every step of $L_1$ corresponds to a (possibly empty) sequence of steps of $L_2$. In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action. The rest of the transitions in $L_1$ are matched to a possibly empty sequence of transitions of $L_2$ with arbitrary labels.

The following lemma shows the soundness and the completeness of $\Gamma$-forward simulations.
\begin{theorem}\label{th:forSim}
Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. If there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$. Moreover, if $L_1$ $\Gamma$-refines $L_2$, $\Gamma=\Sigma_2$, and $L_2$ is deterministic, then there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$.
\end{theorem}
\begin{proof}
TODO
\end{proof}

The linearization of a concurrent history can be also defined in terms of \emph{linearization points}. Informally, a linearization point of 
an operation in an execution is a point in time where the operation is conceptually effectuated; given the linearization points of 
each operation, the linearization of a concurrent history is the sequential history which takes operations in order of their linearization points.
For some libraries, the linearization points correspond to a fixed set of actions. For instance, in the case of atomic libraries  
where method bodies are guarded with a global-lock acquisition, the linearization point of every method invocation corresponds to the execution 
of the body. When the linearization points are fixed, we assume that the library is an LTS over an alphabet that includes actions 
$lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$. The action $lin(m,d,k)$ represents the linearization point of the operation $k$ 
which is an invocation of method $m$ returning value $d$.
Let $LP$ denote the set of such actions. We extend the notion of
well-formedness to sequences $@t$ over alphabet $\Sigma$,
such that $C\cup R\cup LP\subseteq \Sigma$, by adding the following constraints:
\begin{itemize}
	\item every linearization point corresponds to an operation that has been called, i.e., TODO
	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
	\item the return value of an operation is fixed by the linearization point, i.e., TODO
\end{itemize}
%The fact that a history $h_1$ is linearizable w.r.t. another sequential history $h_2$ can be stated as: there exists a way of choosing a linearization
%point 
%Formally, a \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by 
%inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that:
%\begin{itemize}
%	\item every linearization point corresponds to an operation that has been called, i.e., TODO
%	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
%	\item the return value of an operation is fixed by the linearization point, i.e., TODO
%\end{itemize}
We assume that libraries produce only well-formed traces. The projection of a library trace over $C\cup R\cup LP$ is called an 
\emph{extended history}. A trace or extended history is called \emph{LP-complete} when every completed operation has a linearization 
point, i.e., TODO. A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup LP\subseteq \Sigma$ 
and every trace $@t\in Tr(L)$ is $LP$-complete.

Proving the correctness of an implementation $L_1$ of a concurrent object such as a queue or a stack with fixed linearization points
reduces to proving that $L_1$ is a $(C\cup R\cup LP)$-refinement of an abstract implementation $L_2$ of the same object where method
bodies are guarded with a global-lock acquisition. Assuming that the abstract implementation is deterministic, which is typically the case,
by Theorem~\ref{th:forSim}, proving $(C\cup R\cup LP)$-refinement is equivalent to finding a $(C\cup R\cup LP)$-forward simulation 
from $L_1$ to $L_2$~\footnote{Note that the alphabet of $L_2$ is exactly $C\cup R\cup LP$.}.

The next sections extend this result to implementations of queues and stacks where the linearization point of the methods 
that \emph{add} values to the collection is \emph{not} fixed.