### 2017-01-13 draftCAV/draft.tex
started the push implementation

% Suha's Proposal
\documentclass[orivec]{llncs}

\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{arydshln}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{color}

\usepackage{amsfonts}
\usepackage{xspace}
\usepackage{latexsym}
\usepackage{wasysym} % causing problems with llncs's bold vectors
\usepackage{stmaryrd}
\usepackage{alltt}
\usepackage{mathpartir}
\usepackage[ligature,reserved,inference]{semantic}
\usepackage[lined]{algorithm2e}
\def\ttat{\mtt{@}} % the at package clobbers this
\usepackage{at}
\usepackage{alltt}

%\newcommand{\xRightarrow}[2][]{\ext@arrow 0359\Rightarrowfill@{#1}{#2}}

\usepackage[numbers]{natbib}

\input macros

\newtheorem{dfn}{Definition}
\newtheorem{lem}{Lemma}

\title{Existence of Forward Simulations for Particular Data Structures}
\begin{document}

\maketitle

\input{prelim}
\input{refProofs}

\input{queue}
%\section{Existence of Forward Simulations for Stack Implementations that have Fixed Pop Linearization Points}
\input{commit}
\input{set}

\bibliographystyle{abbrvnat}
\bibliography{violin-short}

\newpage
\appendix

\input{app-backSim}
\input{app-absImplQueue}

\end{document}