### 2017-01-16 draftCAV/draft.tex
more on queues

% Suha's Proposal
\documentclass[orivec]{llncs}

\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{arydshln}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{color}

\usepackage{amsfonts}
\usepackage{xspace}
\usepackage{latexsym}
\usepackage{wasysym} % causing problems with llncs's bold vectors
\usepackage{stmaryrd}
\usepackage{alltt}
\usepackage{mathpartir}
\usepackage[ligature,reserved,inference]{semantic}
\usepackage[lined]{algorithm2e}
\def\ttat{\mtt{@}} % the at package clobbers this
\usepackage{at}
\usepackage{alltt}
\usepackage{listings}
\usepackage[lined]{algorithm2e}
\usepackage{wrapfig}
%\newcommand{\xRightarrow}[2][]{\ext@arrow 0359\Rightarrowfill@{#1}{#2}}

\usepackage{pslatex}

\usepackage[numbers]{natbib}

\input macros
\input macros-programs

\lstset{
  language={Java},
  basicstyle=\ttfamily\scriptsize
}

\newtheorem{dfn}{Definition}
\newtheorem{lem}{Lemma}

\title{Linearizability Proofs Based On Forward Simulations}
\author{ }
\institute{ } 
\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\input{intro}
\input{prelim}
\input{refProofs}

\input{queue}
%\section{Existence of Forward Simulations for Stack Implementations that have Fixed Pop Linearization Points}
\input{stack}
%\input{set}

\input{related}

\bibliographystyle{abbrvnat}
\bibliography{violin-short}

\newpage
\appendix

\input{app-backSim}
\input{app-absImplQueue}
\input{app-absImplStack}

\end{document}