### 2017-01-21 draftCAV/intro.tex
finished stacks except TS proof and completed queue stuff

%!TEX root = draft.tex
\section{Introduction}

\newpage~\newpage