### 2017-01-25 draftCAV/app-absImplStack.tex
finished stacks

%!TEX root = draft.tex
\section{Proof of Theorem~\ref{th:absImplStack}}\label{app:absImplStack}

We show that $AbsS$ and $AbsS_0$ refine each other. The standard reference implementation $AbsS_0$ is defined exactly as the one for queues, $AbsQ_0$, except that pop linearization points extract values from the beginning of the sequence stored in the state.

\begin{figure} [t]
{\scriptsize
  \centering
  \begin{mathpar}
    \inferrule[call-push]{
      k\not\in dom(cp^0) \\ 
      d\neq {\tt EMPTY}
    }{
      \sigma,in^0,rv^0,cp^0
      \xrightarrow{inv(push,d,k)}
      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
      \sigma,in^0[k\mapsto d], rv^0,cp^0[k\mapsto A_1]
    }\hspace{5mm}
    \inferrule[lin-push]{
      cp^0(k)=A_1
    }{
      \sigma,in^0,rv^0,cp^0
      \xrightarrow{lin(push,d,k)}
      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
      d\cdot\sigma,in^0,rv^0,cp^0[k\mapsto A]
    }\hspace{5mm}

        \inferrule[ret-push]{
      cp^0(k)=A
    }{
      \sigma,in^0,rv^0,cp^0
      \xrightarrow{ret(push,k)}
      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
      \sigma,in^0,rv^0,cp^0[k\mapsto A_2]
    }\hspace{5mm}

    \inferrule[call-pop]{
      k\not\in dom(cp^0) \\ 
    }{
      \sigma,in^0,rv^0,cp^0
      \xrightarrow{inv(pop,k)}
      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
      \sigma,in^0,rv^0,cp^0[k\mapsto R_1]
    }\hspace{5mm}
        \inferrule[lin-pop1]{
      cp^0(k)=R_1 \\
      \sigma = d\cdot \sigma' 
    }{
      \sigma,in^0,rv^0,cp^0
      \xrightarrow{lin(pop,d,k)}
      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
      \sigma',in^0,rv^0[k\mapsto d],cp^0[k\mapsto R_2]
    }\hspace{5mm}

        \inferrule[lin-pop2]{
      cp^0(k)=R_1 \\
      \sigma = \epsilon
    }{
      \sigma,in^0,rv^0,cp^0
      \xrightarrow{lin(pop,{\tt EMPTY},k)}
      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
      \sigma,in^0,rv^0[k\mapsto {\tt EMPTY}],cp^0[k\mapsto R_2]
    }\hspace{5mm}
    \inferrule[ret-pop]{
      cp^0(k)=R_2 \\
      rv^0(k) = d
    }{
      \sigma,in^0,rv^0,cp^0
      \xrightarrow{ret(pop,d,k)}
      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
      \sigma,in^0,rv^0,cp^0[k\mapsto R_3]
    }\hspace{5mm}
          \end{mathpar}
  }
 \vspace{-4mm}
  \caption{The transition relation of $AbsS_0$. 
  %\textcolor{red}{Call-Enq must have $d!= \texttt{EMPTY}$ as a premise. Also lin deq returning empty must be changed as before.}
  }
  \label{fig:transitions:AbsS_0}
\vspace{-2mm}
\end{figure}

Thus, the states of $AbsS_0$ are tuples $\tup{\sigma,in^0,rv^0,cp^0}$ where $\sigma\in\<Vals>^*$ is a sequence of values, $in^0:\<Ops> ~> \<Vals>$ records the input value of a push, $rv^0:\<Ops> ~> \<Vals>$ records the return value of a pop fixed at its linearization point ($~>$ denotes a partial function), and $cp^0:\<Ops> ~> \{A_1,A,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A,A_2$) or pop operation ($R_1,R_2,R_3$).
All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsS_0}. The alphabet of $AbsS$ contains call/return actions and push/pop linearization points.

To prove that $AbsS$ is a refinement of $AbsS_0$ we define a normal $C\cup R$-backward simulation (i.e, a backward simulation as in Definition~\ref{def:back_app}) from $AbsS$ to $AbsS_0$. The reverse is shown using a normal $C\cup R$-forward simulation (i.e, a forward simulation as in Definition~\ref{def:for_app}).

%We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$. However, we observe that, these implementations contain some points in their pop methods that logically removes the element from the pool. We call them commit points. If a method of a library has a fixed linearization  point, it is also a commit point. In this sense, commit points are weaker versions of fixed linearization points. Fixed linearization point of a pop preserves the following properties:
%\begin{itemize}
%\item If a $ret$ comes before a linearization point in the concrete execution, this order is preserved in the linearization of this execution.
%\item If a $lin$ pop comes before another $lin$ pop in the concrete execution, this order is preserved in the linearization of this execution.
%\item If a $lin$ pop comes before an $inv$ in the concrete execution, this order is preserved in the linearization of this execution. 
%\end{itemize} 
%A commit point is weaker than a fixed linearization point in the sense that it does not need to satisfy the first and the second conditions. 
%
%To our intuition, if a pop (remove) method has multiple finite linearization points, commit point is the latest element in this set. We have never observed an implementation in which a pop is linearized after it logically removes the element from the data structure. 
%
%From now on, we will restrict ourselves to stacks, since our example implementation that we will show linearizability of is the time-stamped (TS) stack. However, the notions and the machines we will introduce can be extended to queues easily. 
%
%We fix $\mathcal{M} = \{push, pop\}$ and $\mathcal{D} = \mathbb{N} \cup \{\texttt{EMPTY}\}$. We extend the alphabet $A\Sigma$ for stacks with commit points as $ACS\Sigma = A\Sigma \cup \{com(pop,d,k)|d \in \mathcal{D}, k \in \mathcal{M}\}$. We define cs-refinement and cs-linearizability as we defined q-refinement, q-linearizability in the previous sections. We also change definitions of backward and forward simulation relations for stacks with commit points as we do in the previous sections by replacing linearization points with commit points in this extensions. Lemma 1 of the first section still holds with new simulation relation definitions and cs-linearizability and cs-linearizability implies the original linearizability definition. 
%
%Our road map for this section is as follows: We will first introduce an intermediate stack machine $L_I$ that will be deterministic with respect to the alphabet $ACS\Sigma$. We will show that $L_I$ is equivalent to the standard abstract stack $L_A$ defined in the previous section with respect to the language $A\Sigma$. We show this by first showing that $L_I$ is a refinement of $L_A$ wrt alphabet $A\Sigma$ by finding a backward simulation relation between them and then $L_A$ is a refinement of $L_I$ with respect to alphabet $A\Sigma$ by finding a forward simulation relation between them. Since $L_I$ is deterministic wrt $ACS\Sigma$, if we have an implementation $L_C$ that is a cs-refinement of $L_I$, we can find a forward simulation between them. As an example, we will pick $L_C$ as time-stamped stack and establish a forward simulation relation between it and our $L_I$ machine. 
%
%Let us continue with defining $L_I$ first:
%\begin{itemize}
%\item A state of $L_I$ again consists of a partial strict order and a program counter: $Q_I \subseteq ND \times ED \times (\mathbb{N} \rightarrow Lbl_I)$ where $Lbl_I = \{N, A_0, A_1, R_0, R_1, R_2\}$ is the set of transition labels for the operations. Different from the fixed linearization point case, this time nodes are not triples but 5-tuples of the form $(k,d,st,mc,con)$. First three fields are the same as previous intermediate machines: $k \in \mathbb{N}$ is operation identifier of a push operation, $d \in \mathcal{D}$ is the data value of that push and $st \in \{\texttt{PENDING}, \texttt{CLOSED}\}$ is the current status of the push operation. The fourth field $mc \subset \mathbb{N}$ keeps the operation identifiers of pop operations such that this push was maximally closed when the pop began. A node $n$ is maximally closed in a state $s$ iff $n.st  = \texttt{CLOSED}$ and if there is an edge $n \rightarrow n' \in ed_s$, then $n'.st = \texttt{OPEN}$. The fifth field $con \subset \mathbb{N}$ is the set of operation identifiers of the pop methods that are concurrent with this push i.e. the either this node was open when the pop started or this node is created when the pop was pending.
%\item The transition labels consist of invocation and return actions of both methods and commit action for only pop method. Hence $\Sigma_I = ACS\Sigma$. Number of parameters for all actions common in previous intermediate stack machine and this intermediate stack machine are the same. Commit actions contain the second data field. They are of the form: $com(pop,d,k)$ where $d \in \mathcal{D}$ and $k \in \mathbb{N}$. 
%\item Initial state consists of an empty strict partial order and a function mapping ever operation to $N$: ${q_0}_I = (\emptyset, \emptyset, f_{{q_0}_I})$ where  $f_{{q_0}_I}(i) =N$ for all $i \in \mathbb{N}$.
%\item We define $\delta_I$ less formally, by giving verbal explanations to the transitions, omitting the obvious updates on the $f$ part and not mentioning about the parts of the nodes that does not change:
%\begin{itemize}
%\item $(q, inv(push,d,k),q') \in \delta_I$ iff a new node $n=(k,d,\texttt{PENDING}, \emptyset, con_n)$ is added to $nd_{q'}$, where $con_n = \{ i \in \mathbb{N}| f_q(i) = R_0\}$; $n' \rightarrow n$ will be added to $ed_{q'}$ if $n'$ is a closed node at the state $q$.
%\item $(q, ret(push,k),q') \in \delta_I$ iff either there is a \texttt{PENDING} node $n$ in state $q$ and this node becomes \texttt{CLOSED} in state $q'$ or there is no node with identifier $k$ in $q$ and nothing else than $f$ field changes in $q'$.
%\item $(q, inv(pop,k),q') \in \delta_I$ iff for every open node $n \in nd_q$, $k$ is added to $n.con$ in $q'$ and for every maximally closed node $m \in nd_q$, $k$ is added to $m.mc$ in state $q'$. 
%\item $(q, com(pop,k,d), q') \in \delta_I$ iff there exists a node $n$ in state $q$ such that $n.d = d$ and either $k \in n.con$ or $k \in n.mc$, this node $n$ and all the nodes adjacent to it are removed in the state $q'$, $k$ is removed from all $con$ and $mc$ fields of all nodes in state $q'$ and for all other pop operations $k'$ in $n.mc$ or in $n.con$ and for all states $n' \in nd_q$ such that $n' \rightarrow n \in ed_q$ and for all $n'' \in nd_q$, $n' \rightarrow n'' \in ed_q$ implies $k' \in n''.con$ (implies $k \notin n''.mc$ but it is stronger than this condition: if there are three closed states $p,q,r$ s.t. $p \rightarrow q$, $q \rightarrow r$, $p \rightarrow r$, $k'$ is only in $r.mc$ and we delete $r$, former condition only allows $k' \in q.mc$ whereas the latter one allows $k' \in p.mc$ in addition) we have $k' \in n'.mc$ in the state $q'$. Note that we need to assume data independence to make this action deterministic.
%\item $(q, ret(pop,k,d), q') \in \delta_I$ iff $q=q'$ ignoring the $f$ fields.
%\end{itemize}
%\end{itemize}
%$L_A$ is the same machine defined in the previous section. The common alphabet between $L_A$ and $L_I$ is $A\Sigma$. We will show that they are equivalent in terms of this alphabet.

\begin{lemma} 
$AbsS$ is a refinement of $AbsS_0$.
\end{lemma}
\begin{proof}
We define a normal $C\cup R$-backward simulation $bs$ from $AbsS$ to $AbsS_0$ as follows. Given an $AbsS$ state $s=\tup{O,<,\ell,rv,cp}$ and an $AbsS_0$ state $t=\tup{\sigma,in^0,rv^0,cp^0}$ we have that $(s,t)\in bs$ iff the following hold:
\begin{itemize}
	\item if a pop has committed or respectively, it has returned in $s$, then it had been linearized or respectively, it has returned in $t$, i.e., for every $k$, if $cp(k)\in\{R_2,R_3\}$ then $cp^0(k)=cp(k)$,
	\item a push is completed in $s$ whenever the same is true in $t$, i.e., for every $k$, $cp(k)=A_2$ iff $cp^0(k)=A_2$,
	\item a push is pending in $s$ iff either it is a non-linearized pending push in $t$ or its linearization point has been executed, i.e., for every $k$, $cp(k)=A_1$ iff $cp^0(k)=A_1$ and $in^0(k)$ doesn't occur in $\sigma$, or $cp(k)=A$,
	\item if a pop didn't commit in $s$ then it is pending in $t$ and it may have been linearized, i.e., for every $k$, if $cp(k)=R_1$ then $cp^0(k)\in \{R_1,R_2\}$,
	\item there exists a partial injective function $g: \{k: cp(k)=R_1\} ~> O$ which associates uncommitted pops to pushes in $O$ such that:
	\begin{itemize}
		\item for every $k$, $k\in dom(g)$ iff $g(k)\in be(k)\cup ov(k)$
		\item  the sequence $\sigma$ is the mirror of a linearization of a partial order $(D,\prec)$ where $D$ contains values labeling elements of $O$ except for those in the range of $g$, and all the values corresponding to completed pushes which are not in the range of $g$, i.e., $\ell_1({\tt COMP}(O)\setminus range(g))\subseteq D\subseteq \ell_1(O\setminus range(g))$ ordered according to the happens-before order between the pushes that added them, i.e., $d_1\prec d_2$ if{f} there exists $k_1,k_2$ such that $\ell_1(k_1)=d_1$, $\ell_1(k_2)=d_2$, and $k_1 < k_2$
		\item every pop in the domain of $g$ has been linearized, i.e., for every $k$, $k\in dom(g)$ implies $cp^0(k)=R_2$,
		\item every pop which is not in the domain of $g$ hasn't been linearized, i.e., for every $k$, $k\not\in dom(g)$ implies $cp^0(k)=R_1$,
		\item every push in the range of $g$ has been linearized, i.e., for every $k$, $k\in range(g)$ implies $cp^0(k)=A$,
		\item a pending enqueue from $O$ has been linearized when its value is contained in $\sigma$, i.e., for every $k$, if $\ell_1(k)\in D$ and $\ell_2(k)={\tt PEND}$, then $cp^0(k)=A$.
	\end{itemize}
	\item the return values fixed at pop commit points are the same, i.e., for every $k$, if $rv(k)$ is defined, then $rv(k)=rv^0(k)$,
	\item every pending push has the same input value in both $s$ and $t$, i.e., for every $k$, $\ell_1(k)=in^0(k)$,
\end{itemize}

In the following, we show that indeed $bs$ is a normal $C\cup R$-backward simulation from $AbsS$ to $AbsS_0$:
\begin{itemize}
	\item Let $s\xrightarrow{inv(push,d,k)} s'$ be a transition in $AbsS$ and $(s',t')\in bs$. We consider two cases depending on whether the value $d$ occurs on a position $i$ in the sequence $\sigma$ of $t'$ or not. If it occurs, let $t$ be a $AbsS_0$ state where essentially, the component $\sigma$ is the prefix of the sequence $\sigma$ of $t'$ that contains the first $i-1$ positions (except for some set of pushes that will be defined hereafter, all operations are at the same control point). Let $@t$ be the following $AbsS_0$ trace:
	\begin{align*}
	@t=inv(push,d,k),\ lin(push,d,k),\ lin(push,d_{i+1},k_{i+1}),\ldots,lin(push,d_{n-1},k_{n-1})
	\end{align*}
	where $d_j$ is the value on position $j$ in the sequence $\sigma$ of $t'$ and $n$ is the length of this sequence (we assume that positions are indexed starting from $0$). Let $k_i=k$.
	For every $k_j$ with $i\leq j\leq n-1$, we must have that $cp(k_j)=A$ in $t'$. We take $cp(k_j)=A_1$ in $t$ for every $j\geq i+1$ and $cp$ undefined for $k_i$.
	We have that $t\xrightarrow{@t} t'$ is a valid sequence of transitions of $AbsS_0$ and $(s,t)\in bs$ (the latter can be proved by taking the same function $g$ used in establishing that $(s',t')\in bs$).
	Now, assume that the value $d$ is not in the sequence $\sigma$ of $t'$.  We consider an $AbsS_0$ state $t$ where the component $\sigma$ is the same as the one in $t'$. There are two sub-cases depending on whether there exists a pending pop $k'$ such that $g(k')=k$ when establishing that $(s',t')\in bs$. If it exists, the operations are at the same control point in both $t$ and $t'$ except for the push $k$ for which $cp(k)$ is undefined in $t$, and the the pop $k'$ for which we take $cp(k')=R_0$ in $t$. We have that 
	\begin{align*}
	t\xrightarrow{inv(push,d,k),\ lin(push,d,k)\ lin(pop,d,k')} t'
	\end{align*}
	in $AbsS_0$. If there exists no such pop $k'$, it can be easily seen that there exists $t$ such that $t\xrightarrow{inv(push,d,k)} t'$ and $(s,t)\in bs$.

	\item Let $s\xrightarrow{inv(pop,k)} s'$ be a transition in $AbsS$ and $(s',t')\in bs$. We consider two cases depending on whether in the function $g$ used to relate $s'$ to $t'$ we have that $k\in dom(g)$. In other words, either the newly invoked pop operation $k$ did not linearize yet ($k\not\in dom(g)$) or it linearizes and removes an element inserted by a linearized push ($k\in dom(g)$). The second case also splits into two sub-cases: The value removed by pop $k$ is inserted by a push $k'=g(k)$ that is still pending or the push has returned. We will look at all three cases separately. 
	The easiest one is the first case. There exists some $t$ where essentially the component $\sigma$ is the same as the one in $t'$, such that $t\xrightarrow{inv(pop,k)} t'$ and $(s,t)\in bs$.

	For the first sub-case of the second case, we take an $AbsS_0$ state $t$ where $\sigma_t = d\cdot\sigma_{t'}$ (we use $\sigma_t$ to denote the component $\sigma$ in $t$). It must happen that $cp^0_{t'}(k)=R_2$. The operations are at the same control point in both $t$ and $t'$, except for $k$ in which case $cp^0$ is undefined. We have that $t\xrightarrow{inv(pop,k),\ lin(pop,d,k')} t'$ and $(s,t)\in bs$. The latter holds because essentially, $k'$ is a maximal node in $s'$ (since it is pending).

	For the second sub-case, we define an $AbsS_0$ state $t$ where the sequence $\sigma$ is the minimal prefix of $\sigma_{t'}$ that includes the value $d$ added by $k'$. Let $i$ be the index of this value in $\sigma_{t'}$ and $k_{j}$ with $i<j$ the identifiers of the pushes that added the values following $d$ in $\sigma_{t'}$.  Let $@t$ be the following $AbsS_0$ trace:
	\begin{align*}
	@t=inv(pop,k),\ lin(pop,d,k),\ lin(push,d_{i+1},k_{i+1}),\ldots,lin(push,d_{n-1},k_{n-1})
	\end{align*}
	where $d_j$ is the value on position $j$ in the sequence $\sigma_{t'}$ and $n$ is the length of this sequence. We have that $t\xrightarrow{@t} t'$ is a valid sequence of transitions of $AbsS_0$ and $(s,t)\in bs$. The latter relies on the fact that $k'$ is a greatest completed push in $s$ and all pushes $k_j$ with $j>i$ are pending in $s$.

	\item Let $s\xrightarrow{com(pop,d,k)} s'$ be a transition in $AbsS$ and $(s',t')\in bs$. When this transition results in removing a greatest completed push or a pending push in $s$, then there exists an $AbsS_0$ state $t$ such that $t\xrightarrow{@t} t'$ is a valid sequence of $AbsS_0$ transitions and $(s,t)\in bs$, for some $t$ and $@t$ defined as in the second case of $inv(pop,k)$. When it removes a completed push which is followed by other completed pushes (in the happens-before in $s$), then we pick $t=t'$. We have that $t\xrightarrow{\epsilon} t'$ and $(s,t)\in bs$ (for the latter we must choose a function $g$ such that $g(k)=k'$ where $k'$ is the push removed by the $AbsS$ transition.

	\item Let $s\xrightarrow{ret(push,k)} s'$ be a transition in $AbsS$ and $(s',t')\in bs$. We consider two cases depending on whether the happens-before in $s$ contains push $k$. If it contains push $k$, there are two sub-cases: (1) if its input is present in $\sigma_{t'}$ then there exists an $AbsS_0$ state $t$ such that $t\xrightarrow{ret(push,k)} t'$ is a valid sequence of $AbsS_0$ transitions and $(s,t)\in bs$, and (2) otherwise, we take a state $t$ where essentially, $\sigma_t = d\cdot\sigma_{t'}$ for which we have that $t\xrightarrow{lin(pop,d,k),\ ret(push,k)} t'$ and $(s,t)\in bs$. If the happens-before in $s$ doesn't contain push $k$, then there exists an $AbsS_0$ state $t$ such that $t\xrightarrow{ret(push,k)} t'$.

	\item The case of pop returns $ret(pop,k)$ is trivial. Such transitions are simulated by $ret(pop,k)$ transitions of $AbsS_0$.
\end{itemize}

%TODO FILL IN THE STEPS
%
%
%We will provide a backward simulation relation $bs$ between states of $L_I$ and states of $L_A$. Our relation $bs \subseteq Q_I \times Q_A$ relates state $q=(so_q, f_q)$ to $q' =(s_{q'}, f_{q'}$ iff (i) for all operation identifiers $k \in \mathbb{N}$, if $f_q(k) \in \{N,R_1,R_2\}$, then $f_q(k) = f_{q'}(k)$; if $f_q(k) = A_0$, then $f_{q'}(k) = A_0$ and the data value $d$ associated with this add operation is not inserted to $s_{q'}$ or $f_{q'}(k) = A_1$;  if $f_q(k) = A_1$, then $f_{q'}(k) = A_2$; if $f_q(k) = R_0$, then $f_{q'}(k)=R_0$ or $f_{q'}(k)=R_1$; (ii) Let us call a pop operation pending if $f_{q}(k) = R_0$ and $PP_q$ be set of pending pops. There there exists a function $g: PP_q \rightarrow ND_q \cup \{\texttt{NONE}\}$ such that $k \in g(k).mc$ or $k \in g(k).con$ for all pop operation identifiers $k$ such that $g(k) \neq \texttt{NONE}$ and $g$ is one-to-one if we neglect \texttt{NONE}; $s_{q'}$ is obtained by extending $so_q$ to a total order in which pending nodes may not take place and nodes (open or closed) $n$ such that there exists a pop operation with identifier $k$ so that $g(k)=n$ surely do not take place, (iii) $g(k) = \texttt{NONE}$ implies $f_{q'}(k) = R_0$ and $g(k) \neq \texttt{NONE}$ implies $f_{q'}(k) =R_1$, (iv) if $g_{q'}(k) =n$ and $n$ is a pending node, then $f_{q'}(k') = A_1$ where $k'$ is the operation identifier part of $n$, (v) if there is a pending node with identifier $k$ and it takes part in the linearization, then $f_{q'}(k) = A_1$. 
%
%Now, we will show that $bs$ is a backward simulation relation:
%\begin{itemize}
%\item[$\langle i \rangle$] $bs[{q_0}_I] =\{{q_0}_A\}$
%\item[$\langle ii-a-push \rangle$] Let $(q,inv(push,d,k),q') \in \delta_I$ and $t' \in bs[q']$. We consider two cases: Either the newly added node in $q'$ takes place in the linearization and there exists an index $i$ such that $s_{t'}(i) = d$ or this new node does not exist in the linearization. For the former case construct $s_t = \langle s_{t'}(1), s_{t'}(2),..., s_{t'}(i-1) \rangle$ and $f_t = f_{t'}$ for all the operations except the ones of which data values are linearized as $s_{t'}(j)$,  for $j\geq i$. For those nodes, $f_{t'}(k') = A_1$ whereas we assign $f_t(k') = A_0$ for $j>i$ and $f_t(k) = N$. Let operation identifiers of these nodes be $k_j$ for $j>i$. Then, $t \xrightarrow{\alpha} t'$ holds where $\alpha = inv(push,d,k), lin(push,d,k), lin(push,d_{i+1},k_{i+1}),....$. Moreover, $t \in bs[q]$ since $s_t$ is a valid linearization of $so_q$ using the same $g$ function and omitting more open nodes and $f_t$ obeys the conditions. For the latter case, we pick $s_t = s_{t'}$. We have two subcases:  There exists a pending pop with identifier $k'$ such that $g(k')$ is the new node with identifier $k$ or not. For the first subcase, we pick $f_t = f_{t'}$ except $f_t(k) = N$ and $f_t(k') = R_0$. Then $t xrightarrow{inv(push,d,k),lin(pop,d,k')} t'$ is a path in $L_A$ and $t \in bs[q]$. For the second subcase, we just pick $f_t = f_{t'}$ except $f_t(k) = N$. Then, it is easy to see that $t \xrightarrow{inv(push,d,k)} t'$ holds and $t \in bs[q]$. 
%\item[$\langle ii-a-pop \rangle$] Let $(q, inv(pop,k),q') \in \delta_A$ and $t' \in bs[q']$. We will again consider two cases: When relating $t'$ to $q'$, either $g(k) = \texttt{NONE}$ or $g(k)$ is a node in $q'$. In other words, either the newly invoked pop operation $k$ did not linearize yet or it linearizes and removes an element inserted by a linearized push. The second case also splits into two cases: The element removed by pop $k$ is inserted by a push $k'$ that is still pending or the push has returned. We will look at all three cases separately. The easiest one is the first case. Construct $s_t=s_{t'}$ and $f_t = f_{t'}$ except that $f_t(k) = N$ whereas $f_{t'}(k)= R_0$. One can see that $t \xrightarrow{inv(pop,k)} t'$ is a step in $L_A$ and $t \in bs[q]$. For the first case of the second case, we construct $s_t = s_{t'} \circ \langle d \rangle$ where $d$ is the data of node identifier $k'$ and $f_t = f_{t'}$ except that $f_t(k) = N$ whereas $f_{t'}(k) = R_1$. One can see that $t \xrightarrow{inv(pop,k), lin(pop,d,k)}$ is a valid path in $L_A$. Moreover, $t \in bs[q]$ since $s_t$ is a valid linearization of $so_q$. This is true because the node with identifier $k'$ is a maximal node in $so_q$ and we can add it to the end of linearization of $so_{q'}$. For the second subcase, we obtain $s_t$   from $s_{t'}$ by the following procedure: Let $n$ be the node with identifier $k'$ and $k''$ be the node such that $k' \rightarrow k''$ is an edge in $so_{q'}$, $k''$ takes part in the linearization of $so_{q'}$ to $s_{t'}$ and it has the minimum index $i$ in the $s_{t'}$ among all such nodes. Then, $s_t = \langle s_{t'}(1), s_{t'}(2),..., s_{t'}(i-1), d \rangle$ where $d$ is the data value of node with identifier $k$. Let $k_j$ and $d_j$ be the identifiers and data values of nodes that constitute $s_{q'}(j)$ for $j>i$. Clearly, $t \xrightarrow{\alpha} t'$ is a path in $L_A$ where $\alpha = inv(pop,k), lin(pop,d,k), lin(push, d_i, k_i),...$. Moreover, $t \in bs[q]$ because $s_t$ is a valid linearization of $so_q$. It is true because $so_q = so_{q'}$, $k'$ is a maximally closed node in $so_q$ and all the nodes with identifier $k_j$ ($j>i$) are pending nodes.
%\item[$\langle ii-c \rangle$] Let $(q, com(pop,d,k), q') \in \delta_I$ and $t' \in bs[q']$. We will consider two cases. The first case is commit action removes a maximally closed or an open node. For these cases, we can construct $t$ as in the second case of the previous item (invoke pop case). The same arguments apply for constructing a path between $t$ and $t'$ in $L_A$ and showing that $t \in bs[q]$.  The second case is that commit action removes a non-maximal closed node. This time, pick $t = t'$. Then, $t \xrightarrow{\epsilon} t'$ is the path in $L_A$ and one can show that $t \in bs[q]$ by choosing $g(k)$ as the node that is removed. 
%\item[$\langle ii-b-push \rangle$] Let $(q,ret(push,k),q') \in \delta_I$ and $t' \in bs[q']$. We consider two cases. Either there is a node $n$ with identifier $k$ and data value $d$ in state $q$ or not. For the first case, we consider two subcases. Either this node takes part in linearization or not (if there exists a pop $k'$ such that $g(k')=n$). For the first subcase, we can pick $s_t = s_{t'}$ and $f_q = f_{q'}$ except that $f_q(k) = A_0$.  Then, $t \xrightarrow{ret(push,k)} t'$ is a path in $L_A$. Moreover, $t \in bs[q]$ since $n$ is a maximal node in $q$. For the second subcase, we pick $s_t = s_{t'} \circ \langle d \rangle$ and $f_t = f_{t'}$ except that $f_t(k) = A_1$ and $f_t(k') = R_0$. Then, $t \xrightarrow{lin(pop,d,k'), ret(push,k)} t'$ is a path $L_A$ and $t \in bs[q]$ since $n$ is a maximal open node in $q$. For the second case, we pick $s_t = s_{t'}$ and $f_t = f_{t'}$ except that $f_t(k) = A_1$. Then, $t \xrightarrow{ret(push,k)} t'$ is a path in $L_A$ and $t \in bs[q]$. 
%\item[$\langle ii-b-pop \rangle$] Let $(q,ret(pop,d,k),q') \in \delta_I$ and $t' \in bs[q']$. We pick $s_t = s_{t'}$ and $f_t = f_{t'}$ except that $f_t(k) =R_1$. One can see that $t \xrightarrow{ret(pop,d,k)} t'$ is a valid action in $L_A$ and $t \in bs[q]$.
%\end{itemize}
\end{proof}

\begin{lemma} 
$AbsS_0$ is a refinement of $AbsS$.
\end{lemma}
\begin{proof}
We define a normal $C\cup R$-forward simulation $fs$ from $AbsS_0$ to $AbsS$ as follows. Given an $AbsS_0$ state $t=\tup{\sigma,in^0,rv^0,cp^0}$ and an $AbsS$ state $s=\tup{O,<,\ell,rv,cp}$ we have that $(t,s)\in fs$ iff the following hold:
\begin{enumerate}
	\item every pop is at the same control point in both $t$ and $s$, i.e., for every $k$ and $i\in \{1,2,3\}$, $cp^0(k)=R_i$ iff $cp(k)=R_i$,
	\item a push has been invoked in $t$ whenever it has been invoked in $s$, i.e., for every $k$, $cp^0(k)=A_1$ iff $cp(k)=A_1$,
	\item a push which is linearized in $t$ has been invoked in $s$, i.e., for every $k$, if $cp^0(k)=A$ then $cp(k)=A_0$,
	\item a push is completed in $t$ iff the same holds in $s$, i.e., for every $k$, $cp^0(k)=A_2$ iff $cp(k)=A_2$,
	\item the pair $(O,\ell)$ in $s$ satisfies the following:
	\begin{itemize}
		\item for every $k$, if $in^0(k)=d$, $cp^0(k)\in \{A_1,A\}$, and $d$ occurs in $\sigma$, then $k\in O$ and $\ell(k)=(d,{\tt PEND})$,
		\item for every $k$, if $in^0(k)=d$, $cp^0(k) = A_2$, and $d$ occurs in $\sigma$, then $k\in O$ and $\ell(k)=(d,{\tt COMP})$,
	\end{itemize}
	\item every pending push in $O$ is overlapping with every non-linearized pop, i.e., for every $k$, if $cp^0(k)=R_1$ then $\{k': k'\in O\land \ell_2(k')={\tt PEND}\}\subseteq ov(k)$.
	\item every completed push is either overlapping or was the greatest completed push before a non-linearized pop started, i.e., for every $k$, if $cp^0(k)=R_1$, then ${\tt COMP}(O)\subseteq ov(k)\cup be(k)$,
	\item for every push that overlaps with a pop $k$ or was maximal in $<$ when $k$ started, its successors are overlapping with $k$, i.e., $k_1\in be(k)\cup ov(k)$ and $k_1 < k_2$ implies $k_2 \in ov(k)$ for each $k, k_1, k_2$
	\item predecessors of pushes in $be(k)$ for a given pop $k$ are neither overlapping with $k$ nor in $be(k)$, i.e., $k_1 < k_2$ and $k_2\in be(k)$ implies $k_1\not\in ov(k)\cup be(k)$ for each $k,k_1,k_2$
	\item pending pushes are maximal in $<$, for every $k$ and $k'$, $k \not< k'$ if $\ell_2(k)={\tt PEND}$,
	\item\label{item:stack_fs_lin} the sequence $\sigma$ is the mirror of a linearization of a partial order $(D,\prec)$ where $D$ contains values labeling elements of $O$ and all the values corresponding to completed pushes, i.e., $\ell_1({\tt COMP}(O))\subseteq D\subseteq \ell_1(O)$ ordered according to the happens-before order between the pushes that added them, i.e., $d_1\prec d_2$ if{f} there exists $k_1,k_2$ such that $\ell_1(k_1)=d_1$, $\ell_1(k_2)=d_2$, and $k_1 < k_2$.
	\item the return values fixed at pop linearization/commit points are the same, i.e., for every $k$, $rv(k)=rv^0(k)$,
	\item every pending push has the same input value in both $s$ and $t$, i.e., for every $k$, $\ell_1(k)=in^0(k)$,
\end{enumerate}

In the following, we show that indeed $fs$ is a normal $C\cup R$-backward simulation from $AbsS_0$ to $AbsS$:
\begin{itemize}
	\item Let $t\xrightarrow{inv(push,d,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$. We have that $(t',s')\in fs$ where $s\xrightarrow{inv(push,d,k)} s'$ (recall that $AbsS$ is deterministic). Since the push $k$ is non-linearized in $t'$, the component $\sigma$ of both $t$ and $t'$ are the same and $\ell_2(k)={\tt PEND}$ in $s'$. Then, the component $\sigma$ in $Abs_0$ states related by $fs$ to $s'$ is allowed to exclude values added by pushes in $s'$ which are labeled as pending. The effect of $inv(push,d,k)$ in $AbsS$ implies that $k$ overlaps with all pending pops.
	\item Let $t\xrightarrow{inv(pop,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$. We have that $(t',s')\in fs$ where $s\xrightarrow{inv(pop,k)} s'$. The only difference between $s$ and $s'$ is that the components $be(k)$ and $ov(k)$ in $s'$ contain the greatest completed pushes in $s$ and the pending pushes in $s$, respectively (these components were undefined in $s$). The relation $fs$ doesn't exclude this particular choice for $be(k)$ and $ov(k)$ when applied to $t'$ and $s'$.
	\item Let $t\xrightarrow{lin(push,d,k)} t'$ be a transition in $Abs_0$ and $(t,s)\in fs$. We have that $(t',s)\in fs$, i.e., the $AbsS_0$ transition is simulated by an empty sequence of $AbsS$ transitions, because essentially the component $\sigma$ of $t'$ still corresponds to a linearization of the pushes in $s$ according to item \ref{item:stack_fs_lin} in the definition of $fs$. The sequence $\sigma$ in $t'$ contains the value added by the push $k$ at the end, but this is allowed by $fs$ since $k$ is labeled as pending in $s$.
	\item Let $t\xrightarrow{lin(pop,d,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$. We have that $(t',s')\in fs$ where $s\xrightarrow{com(pop,d,k)} s'$. The transition labeled by $com(pop,d,k)$ is enabled in $AbsS_0$ because $d$ was the first value in the sequence $\sigma$ of $t$. Indeed, this implies that $d$ was added by a push $k'$ which is maximal in the happens-before stored in $s$. This clearly implies that $k'\in be(k)\cup ov(k)$. In addition, the sequence $\sigma$ in $t'$ does correspond to a linearization of the pushes in $s'$ (which don't contain $k$ anymore) because $\sigma$ in $t$ had this property with respect to $s$ and $\sigma$ in $t'$ is obtained by deleting the first value in the sequence $\sigma$ of $t$. 
	\item Let $t\xrightarrow{ret(push,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$. We have that $(t',s')\in fs$ where $s\xrightarrow{ret(push,k)} s'$. There are two cases depending on whether the value added by $k$ is still present in the sequence $\sigma$ of $t$. If it is not, then the push $k$ doesn't occur in the happens-before from $s$, and the only effect of these two transitions is changing the control point of $k$. Therefore, $(t',s')\in fs$ clearly holds. When this value is still present, the effect of $ret(push,k)$ in $AbsS$ is changing the flag of push $k$ from ${\tt PEND}$ to ${\tt COMP}$. Since the order between pushes doesn't change, we have that $(t',s')\in fs$.
	\item Let $t\xrightarrow{ret(pop,d,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$. We have that $(t',s')\in fs$ where $s\xrightarrow{ret(pop,d,k)} s'$. This case is obvious, the only change between $s$ and $s'$ being the control point of $k$.
\end{itemize}

%We will construct a forward simulation relation $fs$ between $L_A$ and $L_I$. Our relation $fs \subseteq Q_A \times Q_I$ relates state $q = (s_q, f_q) \in Q_A$ to a state $q' =(so_{q'}, f_{q'}) \in Q_I$ iff (i) for all operation identifiers $k \in \mathbb{N}$, if $f_q(k) \in \{N, A_0, R_0, R_1, R_2\}$ then $f_{q'}(k) = f_q(k)$; if $f_q(k) = A_1$, then $f_{q'}(k) = A_0$; if $f_q(k) = A_2$, then $f_{q'}(k)= A_1$; (ii) We form nodes of $so_{q'}$ ($ND_{q'}$ as follows:  If $k$ is a push operation adding data value $d$ and either $f_q(k)= R_0$ or $f_q(k)=R_1$ and data added by this push exists in $s_q$, then there is a \texttt{PENDING} node in $ND_{q'}$ with identifier $k$ and data value $d$. If $f_q(k)= R_2$, then there is a \texttt{CLOSED} node in $nd_q$ with identifier $k$ and data value $d$. If there is an operation identifier $k$ such that $f_q(k) = R_0$, we call this a pending pop and this pop takes place $mc$ or $con$ fields of nodes of $q'$. If $n \in ND_{q'}$ is a \texttt{PENDING} node, then $k \in n.con$. If $n$ is maximally closed, then $k \in n.con$ or $n.mc$. If $k \in n.mc$ or $k \in n.con$ and there exists another node $n' \in ND_{q'}$ such that $n \rightarrow n' \in ED_{q'}$, then $k \in n'.con$. If $k \in n.mc$ and there exists another node $n' \in ND_{q'}$ such that $n' \rightarrow n \in ED_{q'}$, then neither $k \in n'.mc$ nor $k \in n'.con$. For any node $n \in ND_{q'}$, either $k \in n.mc$ or $k \in n.con$. (iii) We form edges of $so_{q'}$ ($ED_{q'}$) as follows: \texttt{PENDING} nodes are maximal. Edges obey the strict partial order conditions. (iv) We can find a linearization $so_{q'}$ that is equal to $s_q$. \texttt{PENDING} nodes may not participate in the linearization. Note that we do not need $g$ function for keeping track of linearized pops unlike the previous proof.
%
%Next, we will show that $fs$ is a forward simulation relation.
%\begin{itemize}
%\item[$\langle i \rangle$] $fs[{q_0}_A] = \{{q_0}_I\}
%$ 
%\item[$\langle ii-a-push \rangle$] Let $(q, inv(push,d,k), q') \in \delta_A$ and $t \in fs[q]$. Pick $t'$ such that $(t, inv(push,d,k), t') \in \delta_I$. Since $s_q=s_{q'}$ and $so_{t'}$ contains a maximal open new node as the only difference from $so_t$, we can linearize $so_{t'}$ so that linearization is equal to $s_{q'}$. By checking the other conditions, one can observe that $t' \in fs[q']$.
%\item[$\langle ii-a-pop \rangle$] Let $(q, inv(pop,k), q') \in \delta_A$ and $t \in fs[q]$. Pick $t'$ such that $(t, inv(pop,k), t') \in \delta_I$. Only difference between $t$ and $t'$ is that maximally closed and open nodes in $t'$ contain $k$ in their $mc$ or $con$ fields. Since this new addition obeys our forward simulation definition, $t' \in fs[q']$.
%\item[$\langle ii-c-push \rangle$] Let $(q, lin(push,d,k) q') \in \delta_A$ and $t \in fs[q]$. Pick $t'=t$. $s_{q'}$ is still a linearization of $so_t$ since the node with identifier $k$ is a maximal node in $ND_t$ and we can linearize it at the end. So, $t' \in fs[q']$ holds.
%\item[$\langle ii-c-pop \rangle$] Let $(q, lin(pop,d,k), q') \in \delta_A$ and $t \in fs[q]$. Pick $t'$ such that $(t, com(pop,d,k) t') \in \delta_I$. The action $com(pop,d,k)$ is a valid action in $L_I$ because $d$ is the last element in $s_q$. Hence, the node $n$ with identifier $k$ and data value $d$ is a maximal element in $so_q$ and either $k \in n.mc$ or $k \in n.con$ by the properties of $fs$. Hence, the node with identifier $k$ can be removed by a $com$ action. In addition, $s_{q'}$ is a linearization of $so_{t'}$ because removed node is a maximal node and $s_{q'}$ is obtained from $s_q$ by deleting the maximum node. Hence, $t' \in fs[q']$ holds.
%\item[$\langle ii-b-push \rangle$] Let $(q, ret(push,k), q') \in \delta_A$ and $t \in fs[q]$. We will consider two cases: either the element inserted by the push with identifier $k$ is removed by a concurrent pop or the element is still in $s_q$. For the former case, there is no node with identifier $k$  in state $t$ and we can pick $t'$ such that $(t, ret(push,k), t') \in \delta_I$. We have $so_t = so_{t'}$ for this case. Since $s_q = s_{q'}$ also holds, $s_{t'}$ is a linearization of $so_{q'}$ and $t' \in fs[q']$. For the latter case, we again pick $t'$ such that $(t, ret(push,k), t') \in \delta_I$ holds. This time, only difference between $so_t$ and $so_{t'}$ is that the node with identifier $k$ is \texttt{CLOSED} in $so_{t'}$. Since the edges are the same, $s_{q'}$ is a valid linearization of $s_{t'}$ and $t' \in fs[q']$ holds. 
%\item[$langle ii-b-pop \rangle$] Let $(q, ret(pop,d,k), q') \in \delta_A$ and $t \in fs[q]$. We pick $t'$ such that $(t, ret(pop,d,k) t') \in \delta_I$. Only difference between $t$ and $t'$ is that $f_{t'}(k) = R_2$ whereas $f_t(k) = R_1$. Since $s_q = s_{q'}$ and $so_t = so_{t'}$, $s_{q'}$ is a valid linearization of $so_{t'}$. By checking the other conditions, we see that $t' \in fs[q']$.  
%\end{itemize}
\end{proof}